### 2022-11-30 paper.Rmd
add paper shell

---
title: "Cookies!"
author: "Benjamin Agyei, Christian Miller, Elliot Shannon, Grayson White"
date: '2022-12-07'
output: pdf_document
fontsize: 12pt
geometry: margin=1in
spacing: double
fig_caption: yes
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gglm)
```

## Introduction

## Methods

## Results

## Discussion

## Conclusion